Is Partner Modeling Mutual?

### 2014-08-27 project-mutual-modeling.tex
Two columns, figure layout
Raw import of studies 2->5 from Word

\documentclass{article}

@@

\begin{figure*}[htb]
\centering

@@

\label{mm_symmetry}
\end{figure*}

@@

\label{mm_rectangle}
\end{figure*}

@@

\subsection{Study 1: Effect of an awareness tool on \Model{A}{B} in a virtual
game}

@@

\begin{figure}
        \centering
        \includegraphics[width=0.9\columnwidth]{image4.png}
        \caption{Snapshot of the SpaceMiner Game}
        \label{spaceminer}
\end{figure}

@@

Figure 4: M(A,A) and M(A,B) questionnaires in SpaceMiners (translated from French)

@@

\bibliographystyle{abbrv}
\bibliography{biblio}

>>>>

\documentclass[twocolumn]{article}

@@

\begin{figure}[htb]
\centering

@@

\label{mm_symmetry}
\end{figure}

@@

\label{mm_rectangle}
\end{figure}

@@

\subsection{Study 1: Effect of an awareness tool on \Model{A}{B}{x} in a virtual
game}

@@

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{image4.png}
        \caption{Snapshot of the SpaceMiner Game}
        \label{spaceminer}
\end{figure*}

@@

\begin{figure}[ht!]
        \centering
        \includegraphics[width=\columnwidth]{image5.png}
        \caption{M(A,A) and M(A,B) questionnaires in SpaceMiners (translated
        from French)}
        \label{study1:questionnaires}
\end{figure}

@@

\subsection{Study 2: Effect of an awareness tool on M°(A,B)  in a pervasive
game}

This study concerns a collaborative game that occurred in the physical space
(Nova et al , 200?). We studied whether players build an accurate model of the
path followed by their partners because this path reflected their problem
solving strategy. We used an objective measure of M°(A,B): the distance between
where A believes B has been walking and where B actually went. The main
hypothesis concerned the effect of awareness tools on group performance and on
M°(A,B).

CatchBob is a mobile game in which groups of 3 players have to solve a joint
task. The game was played on EPFL campus. Participants had to find a virtual
object ('Bob') and "to catch it" by forming a triangle around it. Players used a
Tablet PC that displayed a map of the campus and an indication of their personal
distance from Bob. Their annotations on the map were shared with the two other
players (A could see what B and C wrote). These annotations faded out after a
few minutes to avoid covering the full display. The awareness tool displayed the
location of the two other players on the map. The availability of location
awareness tool constituted our independent variable.

\begin{figure*}[h!t]
        \centering
        \begin{subfigure}{.3\textwidth}
            \includegraphics[width=\linewidth]{image6.png}
            \caption{A's drawing A of B's path.}
        \end{subfigure}
        \begin{subfigure}{.3\textwidth}
            \includegraphics[width=\linewidth]{image7.png}
            \caption{Actual path followed by B.}
        \end{subfigure}
        \caption{Bla bla}
        \label{study2:paths}
\end{figure*}

Ninety students participated in this experiment. We chosen only students from
our University because knowing the campus geography had an impact both on group
performance but also on mutual modeling: it is hard to represent the path of
someone across some space without having a mental map of this space. We selected
groups of students who knew each other. We assigned 10 triads to each of our
three experimental conditions. Each condition was made up of approximately 25%
of women, but we did not control gender repartition within each triad.

The independent variable was the awareness tool: the control condition (without
tool) and two experimental conditions: synchronous awareness (display current
position of each player) and asynchronous awareness (display current position of
each player and their recent spatial trace).  As a dependent variable, we had
the task performance which was the distance covered by the team to catch Bob and
M°(A,B). To estimate M°(A,B), we asked players to draw on paper their own path
and the path of each of their partners after the game. This enabled us to
calculate the number of errors players made while drawing the path of their
partners. We compared the path that player A attributed to B with B's real paths
recorded by the system and the same for A\&C or B\&C as depicted on Figure 5.

M°(A,B) is the sum of errors made by A about B’s paths. An error was either
drawing a place where the partner had not been or not drawing a place where
he/she had gone. One could argue that M°(A,B) is biased by the subjects' ability
to translate their trajectories souvenir into a map drawing. However, 85% of
subjects made no mistake at all when drawing their own path. We therefore
consider mistakes in their partners' path as being due to a lack of mutual
modeling accuracy instead of being due to spatial reasoning skills.

Grounding criterion. The correlation between M°(A,B) and the task performance
(path lengths) was low: 0.15. Using a post-hoc split on M°(A,B), there is no
significant difference between the performance of the groups with high and low
M°(A,B)  (F = 1.45, p = 0.24). Conversely, a post-hoc split of the groups
according to their performance did not show any significant differences on
M°(A,B) (F = 1.16, p = 0.29).

Study-specific questions. There was no significant difference regarding the task
performance. However, our surprise was that the absence of the awareness tool
was related to a higher M°(A,B): players better remembered their partners’ path
if they did not see their position permanently! We will not enter into the
details of these results but, to be short, teams without awareness tool made
more annotations on the map while permanent MLA has an underwhelming effect
(Nova et al., 2005).

Symmetry question. The correlation between M°(A,B)  and M°(B,A)  is positive (r
= .41) and significant (p < .01);  the more A makes errors about B, the more B
does the same.

Triangle questions. About ∆2, the correlation between M°(A,B) and M°(A,C)is
significant: r=0.30, p <.01. About ∆3, the correlation between M°(A,C) and
M°(B,C) is significant: r=0.43, p <.001.

This positive result regarding to the symmetry question is even more surprising
than the previous one: despite the high heterogeneity of spatial skills among
adults (Liben et al, 1981), a significant correlation could indicate that mutual
modeling relies more on group processes than on individual processes. The
significant results on the triangle questions probably go in the same direction:
either the modeler or the modeled both contribute to accurate modeling, or that
the quality of teamwork has a global effect. In this experiment, since team
members did not interact intensively during the task, the factor that
facilitates accurate mutual modeling is probably less the quality of verbal
interactions that the group strategy: a clear strategy facilitates memorizing
one’s partner path.

\subsection{Study 3:  Effect of media richness on M°(A,B) in argumentation}

The aim of this unpublished study was to evaluate the effect of media richness
on M°(A,B,emotions). The hypothesis was that video communication would lead to a
better M°(A,B) than audio only since emotions often impact on facial
expressions.

\begin{figure*}[ht!]
        \centering
        \includegraphics[width=0.9\textwidth]{image8.jpg}
        \caption{Example of argumentation graph}
        \label{study3:argumentation_graph}
\end{figure*}

\paragraph{Experimental settings}

Triads had to address an emotional societal debate: authorizing or not adoption
by homosexual couples). They worked on-line and had to structure their
argumentation with a shared concept map tool (TeamWave) as illustrated in Fig.
6. Ten groups had only an audio connection while ten groups had audio plus
video. The video communication was provided by a web cam and the software
IVisit. For the audio link, we used microphones, headsets and a software called
BattleCom. In the audio+video condition, the screen was divided in three
sections. The main part was devoted to the concept map window (24 X 24 cm). In
the remaining part appeared the image of the two peers (8.5 X 8.5 cm each). In
the audio condition, this video zone was left empty so that the size of the
concept map was equal in each condition. The subjects were located in the same
room, separated by mobile walls. Despite their headsets, it is not impossible
that they could hear non-verbal audio cues (e.g. tapping the sole with feet).
The task last in average 61 minutes.

Twenty triads from the University of Geneva participated to this experiment. We
formed groups of subjects who knew each other, i.e. they had ab initio mutual
models: the task required the discussion of sensitive issues which required to
feel quite comfortable with peers. There were 36 women and 24 men, but, since
groups were formed a priori, we could not control the balance of men and women
in each condition.

The independent variable was the presence of not of video link. The dependent
variable M°(A,B) was subjectively by using 3 questionnaires: in the first one, A
described his or her own emotions M(A,A) , while in the two other questionnaire,
A describes B’s and C’s emotions.   The questionnaire included 18 items (7 steps
Likert Scale) describing emotions labeled as adjectives: anxious, enthusiastic,
agitated, proud, excited, quiet, calm, stressed, bored, upset, relaxed,
irritated, determined, hostile, active, etc. This list of emotions was
elaborated from (REF?). The proposed emotions were . M(A,B) is a vector of 18
numerical values {x1 … xi} corresponding to their answers on each questionnaire.
M°(A,B) has been computed as the correlation between the two vectors :

M°(A,B,emotions) = ∆M(A,B,{x1 … xi}, M(B,B,{y1, … ,yi}) = r ({x1 … xi}, {y1, …
,yi}))

\begin{figure*}[ht!]
        \centering
        \includegraphics[width=0.9\textwidth]{image9.pdf}
        \caption{Average values of M°(A,B,X) where X is one of the proposed
        emotions (max = 7).}
        \label{study3:deg_m_values}
\end{figure*}

Grounding criterion. The maps produced by teams were ranked. We use the average
rank as a estimation of the team performance and correlated with the average of
the 6 values of M°(A,B) per team (M°(A,B), M°(A,C), M°(BA),…): The correlation
is -0.22: teams with a good M°(A,B) team to be better ranked.

Study-specific questions. Our hypothesis about media richness is rejected: the
average M°(A,B, emotions) was .62 (SD .28) with the audio+video condition and
.71 (SD .19) in the audio condition (F=3.4, p: 0.065). Our interpretation is
that viewing what one's partner sees (shared graphical editor) was, in this
task, more important than seeing each other (Gaver et al 1993; Anderson et al.,
2000).

Symmetry questions. Regarding the main items, the mutual modeling of emotions,
the correlation is .25 or .36  a reprendre avec - Patrick

Triangle questions. Check (∆2) . For (∆3) If we compute for each subject A, the
correlation between the vectors of emotions in M(B,A) and the vector in M(C,A),
and take the average across all 60 subjects, we obtain a very impressive
correlation of .66. This average is about the same in each condition.

Rectangle questions We cannot address the relationship (∆4) between self-and
social accuracy here because we don't have an estimation of self-accuracy:
subjects describe their emotions but we have no way to check if there are
correct. However, by measuring M°(A,B) on  18 emotional labels, we can however
have a glimpse about ∆5: how M°(A,B,X) varies according to X. Figure 7 shows the
range of modelling errors: the difference between M(A,B) and M(B,A), on a scale
of 7, is in average of 0.3 for the emotion "discasted" up to 1.9 for the emotion
"calm". This is probably specific to the variety of scales  (SD= 0-5 for
"discated" versus .7 for "calm"), our point is not to interpret this too far,
but to show that there are large varations even with one area  (perceiving
emotions).

Chek si il faut modifier les resultats de la symmetries

\subsection{Study 4:  Effects of a script on M°(A,B)  in concept mapping}

This study investigated the effect of a collaboration script on collaborative
learning (REF).. The script chosen is a JIGSAW: two students receive different
but complementary subsets of the knowledge (texts) which have to be integrated
to build a shared concept map.  This script increases the cognitive effort to
build the map, not only to recon ciliate the viewpoints of each team members
but, before that, to find out what the other knows.

\paragraph{Experimental settings}

The instructional material consisted of an explanatory text about the
neurophysiologic phenomenon of “action potential”. The text was divided into 3
chapters. The peers were located in two rooms equipped with the same computer.
The experimental session lasted around 90 minutes and consisted of 6 phases:
Participants used two software components, CmapTools (http://cmap.ihmc.us/) and
TeamSpeak (http://www.goteamspeak.com/).

1. As a pre-test, participants were asked to write down all they knew about the
neuron and its functioning (5 minutes).

2. Participants were instructed to read a text (12 minutes).

3. Participants were asked to build individually a concept map in order to
graphically represent what they learnt from the text (10 minutes).

4. Dyads had to built a concept map during 20 minutes, communicating by audio.
The screen layout was structured into three areas (see Figure 8)

5. Participants were invited to individually complete a knowledge test (15-20
minutes).

6. Participants where asked to estimate their own- and their partner’s final
knowledge in a questionnaire.

Figure ?...

\begin{figure*}
    \centering
    \begin{tikzpicture}[scale=0.8]
        \node[anchor=south west,inner sep=0] at (0,0)
            {\includegraphics[width=0.65\textwidth]{image10.png}};

            \node[rectangle,draw, thick] at (11,6.2) {Self map};
            \node[rectangle,draw, thick] at (11,3) {Partner map};
            \node[rectangle,draw, thick] at (3,2) {Collaborative map};
    \end{tikzpicture}
    \caption{The group concept map and the individual concept maps in study
    4}
    \label{study4:concept_map}
\end{figure*}

Fifty-eight first year students from EPFL, 47 men and 11 women, with a mean age
of 20.46 were remunerated for participation. Dyads were randomly assigned to one
of the two experimental conditions. The men/women ratio was equivalent in each
condition. Participants did not know each other before the experiment. Students
from Life Sciences were not recruited since they had too high background
knowledge on the learning domain.

The independent variable, script versus no-script, was implemented by the
difference of texts that individuals had to read. In the “same information” (SI)
condition, the same text was given to both partners. In the “complementary
information” (CI) condition. it was divided into two sub-texts, one about the
electrical processes of the neuron while the second one about the chemical
processes. These two versions were equivalent in terms of number of information
pieces.

The dependent variables were the post-test scores, but this paper is mostly
concerned by the M(A,B, knowledge).  In phase 6, participants were ask to
estimate (7-point Likert scale) their own- and their partner’s outcome knowledge
with respect to each chapter of the zexz. The order of questions about oneself
and about the other was counterbalanced across participants. We measure M°(A,B)
in an objective (1 \& 2) as well as in a subjective way (3).

(1) M°(A,B, knowledge) = Σi=1to3  | M(A,B, chapteri) – score (B, chapteri) |

(2) M°(A,A, knowledge) = Σi=1to3  | M(A,A, chapteri) – score (A, chapteri) |

(3) m°(A,B, knowledge) = Σi=1to3  | M(A,B, chapteri) – M(B,A, chapteri) |

Grounding criterion. In this task, the grounding criterion was low, but we did
not evaluate task performance (e.g. the quality of the jointly produced concept
map) but the learning gains. The correlation between M°(A,B) and A's learning
gains is not significant (β =.08, ,NS, N = 60). It is also not significant
within each condition.

Study-specific questions. We performed a non-parametric Mann-Whitney test on
post-test scores for electric questions and a one-way ANOVA on scores for ionic
questions (Levene tests for homogeneity of variances: p = .02 and ns,
respectively. Results  did not show any significant difference between the “same
information” (SI) condition and the “complementary information” (CI) condition,
neither for electric questions [U = 388.50, z = -0.88, ns] nor for ionic
questions [F(1, 58) = 0.17, ns].  The effect of scripts on M°(A,B) was not
significant (F(1, 58) = 0.78, ns) when considering the absolute difference
between M°(A,B) and B's post-test score. However, A tended underestimate B's
score in the SI condition (M = -2.06) and to overestimate it in the CI condition
(M = 1.21). (F(1, 58) = 6.44, p<.01). Regarding to M°(A,A), there was no
significant difference between conditions.

Symmetry question. The interclass correlation between M°(A,B) and M° (B,A) is
not significant, but almost (r=.26 (F(1,29)=1.71, p= 0.075)). Actually, it is
significant when students read the same text (SI condition: r= 0.43,
F(1,15)=2.53, p<.05) but not when they read different texts (r=.13, F(1,12)=1.3,
NS).

Triangle questions do not apply here since we worked with dyads.

Rectangle question ∆4: the correlation between M(A,A) and M(A,B) is globally not
significant (r=0.05) as well as not significant in each condition: someone good
at self-modeling is not necessarily good at modeling someone else and
vice-versa.  Discussion

In this experiment, the symmetry of mutual modelling is only found when subjects
are in conditions of symmetry of information. This consolidates the hypothesis
of M°(A,B) being a property of interactions in context rather than A's cognitive
attitude.

\subsection{Study 5:  M°(A,B, knowledge)  in concept mapping}

This study investigates if M°(A,B, knowledge) is related to learning outcomes by
comparing teams with or without a knowledge awareness tool (KAT), i.e. a tool
that informs A about B's knowledge as measured through a pre-test.

The peers were located into two different rooms. A complete description of the
study is provided in Sangin et al (2008).The experiment lasted 90 minutes. It
started with the same two steps as in study 4.

3. Subjects passed a pre-test, with ten questions per chapter.

4. Participants had 20 minutes to draw a collaborative concept map reporting the
content of the texts. They were able to communicate orally through headsets.  We
used Tobbii eye tracking devices to record their gazes.

5. The post-test included the same items than the pre-test but in a different
order.

6. Finally, participants were asked to estimate their partner’s knowledge at the
post-test for each of the three chapters on a 7-point likert-like survey.

\begin{figure*}
        \centering
        \includegraphics[width=0.9\textwidth]{image11.png}
        \caption{Screenshot of the KAT condition during the concept-map building
        phase.}
        \label{study5:kat}
\end{figure*}

Sixty-four first year EPFL students (18 women, mean age = 21.2 years)
participated to the study. They were involved in curricula that do not imply
advanced neurophysiologic notions. They were randomly assigned to conditions and
did not know each other before.

The participants of the experimental condition group were provided with the KAT
on the bottom part of the screen (see Figure 9): each line represents the score
obtained by the partner at the pre-test for a chapter. Participants did not see
they own score, but they usually started their discussion by exchanging this
information.

Grounding criterion. A linear regression revealed a positive relation between M°
(A,B) and the collaborative learning gain of the pair (AB): β = .401, p < .001,
r2adj. =.15, large effect. This relationship was not significant in the previous
experiment which as been conducted with the same task: using the same task
should produce the same grounding criterion This probably due to the fact that
M° (A,B) was influence by the KAT.

Study-specific questions. The t-test reported a significant difference between
the KAT condition participants (M = 13.4\%) and the control group [M = 3.6%, t(1,
60) = 2.73, p < .01,Cohen’s d = 0.7, medium to large effect]. Providing learners
with cues about the prior-knowledge of their partner enhances their
collaborative learning. As a treatment check, we found a positive and
significant correlation between the amount of gazes on KAT (using eye tracking
devices) and the learning gains (r(22) = .54, p = .01). A detailed analysis
revealed that subjects look at KAT to assess their peer's credibility when (s)he
provided new information.

The KAT has a significant effect on M°(A,B): peers more accurately estimated
their partners knowledge (M = 1.11) than those in the control condition [M =
.98, t(1, 60) = 3.19, p < .01, Cohen’s d = 0.83, large effect]. This is a
trivial result since the KAT provided them with an initial M(A,B). However,
subjects have to predict the post-test score while the KAT informed them about
the pre-test score. Actually, pairs in the KAT condition produced significantly
more instances of 3 specific categories of interactions: (1) elaborated
utterances with rich contents,  (2) utterances asking about the other's
knowledge such as "Did you understand how transmission works?" (3) utterances
describing one's own knowledge (M(A,A)) such as "I don’t remember the Ranvier’s
thing…" . These 3 categories provide different account of M°(A,B): (1) as an
effect of the quality of interaction, (2) as an effect of A's effort to model B
or (3) as B's effort to give cues to A about his own knowledge.

We examined M° (A,B) as potentially mediating the effect of the KAT factor on
the relative learning gain. A linear regression confirmed that M° (A,B) was
significantly related to the KAT-factor (β = .381, p < .01, r2 = .15). The
KAT-factor was also significantly and positively related to the RLG (β = .332, p
< .01, r2 = .11). We then tested the relation between the independent variable
(KAT) and the dependent variable (gains) when controlling for the mediating
variable (M° (A,B)). A multiple regression showed that the KAT-factor was no
longer a significant predictor (β = .210; p = ns) whereas the M° (A,B) was still
a significant predictor (β = .32, p < .01). Thus, it can be concluded that M°
(A,B) mediated the KAT-factor’s effect on the learners’ RLG. The Sobel (1982)
significance test for indirect effects was significant [z = 1.99, p < .05].

Symmetry question.  We did not find an intra-pair correlation (ICC = 0.05, NS).
Rectangle question ∆4: the correlation between M(A,A) and M(A,B) is \fixme{???
MIRWEIS A FAIRE}.

Despite the fact that the learning task was the same as in study 4, the
conditions of collaboration (viewing multiple maps or not) and the conditions
(scripted or not, awareness tool or not) probably explain differences in terms
of mutual modeling

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliographystyle{abbrv}
\bibliography{biblio}
